### 2019-08-19 abstract.tex
Final draft.

\begin{resumen}

\end{resumen}

%\begin{abstract}

%\end{abstract}

>>>>

%\begin{resumen}

%\end{resumen}

\begin{abstract}
The Transit Monitoring in the South (TraMoS) project aims to perform photometric follow-up in known transiting exoplanets. This process provides not only essential information about the transiting planet but also about their planetary system's architecture. Departures from the linear ephemeris given by Keplerian motion, the so-called Transit Timing Variations (TTV), can only be spotted through extensive follow-up, and could suggest the existence of additional bodies in the system, such as planetary companions or exomoons.

This Master thesis recapitulates the last years of the TraMoS project, on which new targets were selected, modern tools were used and new analysis were performed. I studied 27 new light curves of three hot Jupiters: WASP-18b, WASP-19b, and WASP-77Ab. Their orbital and physical parameters were refined, and their linear ephemeris updated thanks to the inclusion of archival transit time data. Given the lack of significant variations in their transit times, dynamical analysis can place upper mass limits in hypothetical pertubers. Bodies with masses greater than $50\,M_{\odot}$ can be discarded for WASP-18b, $6\,M_{\odot}$ for WASP-19b, and $10\,M_{\odot}$ for WASP-77Ab. Furthermore, I included recent TESS data for the three targets, supporting the linear ephemeris as the best fit for their transit times

The results presented in this thesis support the rare occurrence  of companions in hot Jupiters systems. Further studies may confirm if this kind of planet goes through a different formation process ending up orbiting alone on their systems or they have very small and unseen companions.
\end{abstract}

### 2019-08-19 conclusion.tex
Final draft.

The TraMos project started more than 10 years ago aiming to refine orbital and physical parameters of already known exoplanets, and to search for variation in the transit time -- or other planet parameters -- that could suggest the existence of smaller bodies on their systems. Since the beginning, the targets have been hot Jupiters due to their short-periods, allowing to obtain more consecutive transit light curves in a reduced timespan.

@@

During this Master project, I studied three exoplanets: WASP-18b, WASP-19b, and WASP-77Ab. For these targets, I analyzed a total of twentysix new light curves: eight of WASP-18b, nine of WASP-19b and nine of WASP-77Ab. In the case of WASP-77Ab, four light curves were taken from the Exoplanet Transit Database (ETD)\footnote{\url{var2.astro.cz/ETD}} to obtain a wider time coverage. For the light curves coming from the TraMoS project, I started from performing aperture photometry of the targets using our data reduction pipeline (for further details see \ref{chap:pipeline}), in order to obtain high-precision light curves. Then, I used EXOFASTv2 \cite{Eastman2013,Eastman2017} to get the best-fit for the light curve's data and hence, to refine their orbital and physical parameters and obtain their transit times.

The TTV curves were constructed from the residuals between the observed transit time and the expected from the linear ephemeris. None of the targets studied in this thesis, shows a clear, periodic sinusoidal signal in the variation of their transit times and moreover, the structure of the transit times are comparable with a random distribution after a Lomb-Scargle analysis. This suggests the lack of close-in companions in MMR in their systems. However, upper mass limit of hypothetical perturbers -- placed not only in near MMR -- could be computed with the time-scatter from the linear ephemeris. The hot Jupiter WASP-18b shows 83 seconds of scatter, WASP-19b shows 75 seconds and WASP-77ab shows 121 seconds. The latter being the target with larger deviation from the linear ephemeris, but at the same time, the target with less transit times.

@@

After the release of TESS data, I included transit times of the targets studied in this thesis, in order to extend their TTV curves and confirm the lack of periodic variations in the transit time. The results came out as expected and the scatter in the transit time were reduced for the three targets: 47 seconds for WASP-18b, 65 seconds for WASP-19b and 86 seconds for WASP-77Ab. Moreover, I refined each orbital period obtaining a precision of the order of milli-seconds.

To date, none of the previous and current targets of the TraMoS project have shown TTVs. As they are all hot Jupiters, these studies support the theory of "alone hot Jupiters", on which this kind of exoplanets are probably alone in their systems or accompanied of small bodies not in MMR. How WASP-47b and Kepler-730b, the only two hot Jupiters showing TTVs, have close-in companions is still unknown. The migration theory of hot Jupiters could give us a possible explanation of how these gas giant ended-up with companions in MMR, as they are supposed to start their migration process in far-away orbits. While migrating, they could catch other bodies to then settling down in close orbits. But the lack of more hot Jupiters with TTVs suggest that probably during migration they cleaned-up their orbits of small bodies.

The TraMoS catalogue stores more than 400 transits events of near 140 exoplanets. As future work I propose to continue performing ground-based photometric follow-up of the targets in this catalogue, as well as new targets, combining the efforts with TESS data of Sector 1 to 13. In this way, we could obtain better estimations of TTVs with wider time-span data from TraMoS, and an important amount of consecutive transit times from TESS. In the near future, I will analyze the exoplanets WASP-36b and HATS-34b, which each one has five transit times from TraMoS and twelve and twenty six from TESS, respectively. For WASP-22, WASP-46, WASP-45, WASP-23 and WASP-123, TESS provided five, seventeen, eight, seven and thirteen light curves, respectively. These targets are within the 30 targets with most light curves from TraMoS, therefore, they should be the next in line.

The reduction pipeline is fully functional and provides high-precision photometry, nevertheless several improvements were left behind because of the lack of time to implement them. For example, the coordinates selection could be done inside the program, to avoid mistakes when typing them by hand before calling the pipeline. Any new idea to improvement the reduction pipeline could be done within an undergrad project or graduate semester-workshop.

>>>>

The TraMos project started more than 10 years ago aiming to refine orbital and physical parameters of already known exoplanets, and to search for variation in the transit time -- or other planet parameters -- that could suggest the existence of smaller bodies on their systems. Since the beginning, the targets have been hot Jupiters due to their short-periods, allowing to obtain more consecutive transit light curves in a reduced period.

@@

During this Master project, I studied three exoplanets: WASP-18b, WASP-19b, and WASP-77Ab. For these targets, I analyzed a total of twenty-six new light curves: eight of WASP-18b, nine of WASP-19b and nine of WASP-77Ab. In the case of WASP-77Ab, four light curves were taken from the Exoplanet Transit Database (ETD)\footnote{\url{var2.astro.cz/ETD}} to obtain a wider time coverage. For the light curves coming from the TraMoS project, I started from performing aperture photometry of the targets using our data reduction pipeline (for further details see \ref{chap:pipeline}), in order to obtain high-precision light curves. Then, I used EXOFASTv2 \cite{Eastman2013,Eastman2017} to get the best fit for the light curve's data and hence, to refine their orbital and physical parameters and obtain their transit times.

The TTV curves were constructed from the residuals between the observed transit time and the expected from the linear ephemeris. None of the targets studied in this thesis shows a clear, periodic sinusoidal signal in the variation of their transit times and moreover, the structure of the transit times are comparable with a random distribution after a Lomb-Scargle analysis. This suggests the lack of close-in companions in MMR in their systems. However, upper mass limits of hypothetical perturbers -- placed not only in near MMR -- could be computed with the time-scatter from the linear ephemeris. The hot Jupiter WASP-18b shows 83 seconds of scatter, WASP-19b shows 75 seconds and WASP-77ab shows 121 seconds. The latter being the target with a larger deviation from the linear ephemeris, but at the same time, the target with fewer transit times.

@@

After the release of TESS data, I included transit times of the targets studied in this thesis, to extend their TTV curves and confirm the lack of periodic variations in the transit time. The results came out as expected and the scatter in the transit time were reduced for the three targets: 47 seconds for WASP-18b, 65 seconds for WASP-19b and 86 seconds for WASP-77Ab. Moreover, I refined each orbital period obtaining a precision of the order of milliseconds.

To date, none of the previous and current targets of the TraMoS project have shown TTVs. As they are all hot Jupiters, these studies support the theory of "alone hot Jupiters", on which this kind of exoplanets are probably alone in their systems or accompanied of small bodies, not in MMR. How WASP-47b and Kepler-730b, the only two hot Jupiters showing TTVs, have close-in companions is still unknown. The migration theory of hot Jupiters could give us a possible explanation of how these gas giant ended-up with companions in MMR, as they are supposed to start their migration process in far-away orbits. While migrating, they could catch other bodies to then settling down in close orbits. But the lack of more hot Jupiters with TTVs suggest that probably during migration they cleaned-up their orbits of small bodies.

The TraMoS catalog stores more than 400 transits events of near 140 exoplanets. As future work, I propose to continue performing a ground-based photometric follow-up of the targets in this catalog, as well as new targets, combining the efforts with TESS data of Sector 1 to 13. In this way, we could obtain better estimations of TTVs with wider time-span data from TraMoS, and an important amount of consecutive transit times from TESS. In the near future, given the new scatter of the transit time in the three targets analyzed in this thesis, more accurate upper mass limits for hypothetical perturbers could be computed. Also, I will analyze the exoplanets WASP-36b and HATS-34b, which each one has five transit times from TraMoS and twelve and twenty-six from TESS, respectively. For WASP-22, WASP-46, WASP-45, WASP-23, and WASP-123, TESS provided five, seventeen, eight, seven and thirteen light curves, respectively. These targets are within the 30 targets with most light curves from TraMoS, therefore, they should be the next in line.

The reduction pipeline is fully functional and provides high-precision photometry, nevertheless, several improvements were left behind because of the lack of time to implement them. For example, the coordinates selection could be done inside the program, to avoid mistakes when typing them by hand before calling the pipeline. Any new idea to improve the reduction pipeline could be done within an undergrad project or graduate semester-workshop.

### 2019-08-19 introduction.tex
Final draft.

Having billions of billions of galaxies in the Universe each one with billions of stars, the current number of discovered exoplanets is still very low. Furthermore, the hunting zone of exoplanets is limited by our current technology and finding an exoplanet in the closer galaxy -- Andromeda -- can be considered science fiction. If the Milky Way has \textit{only} around $2\,\times10^{11}$ stars, how could be the Solar System unique? How could be the Earth the only planet with life on it? Until today, it is still difficult to answer those questions. The Solar System presents a logic structure where the small rocky planets are closer to the Sun, and the gas giants are located beyond. But no other exoplanetary system have been found having similar features and even not exoplanetary life has been detected yet\footnote{By the date of writing this thesis}. Our knowledge about the Solar System challenges the existence of other kinds of planetary architectures with close-in gas giants, for instance, suggesting that there is not a single, universal planetary process formation. Answering such questions makes studying exoplanets a relevant matter nowadays.

@@

Anyway, the rising research field of exoplanets does not end with their discovery. Performing follow-up and proper characterization of discovered exoplanets is essential to obtain a full list of orbital and planetary parameters. For instance,  39\%  of the total of discovered exoplanets have a proper characterization of their masses, and only 17\% have their mass and radius computed. Moreover, through follow-up studies several phenomena, such as atmospheres, can be unveiled.

Photometric follow-up of transiting exoplanets provides not only fundamental information about the exoplanet, but their environment. Variations in the transit timing -- when the planet passes in front of its star -- could be induce by companions gravitationally bounded, such as other planets in their systems or even exomoons. As most of the space mission have strict fields of view, performing photometric follow-up with ground-based telescopes is crucial to reveal important features of their systems. The Transit Timing Variation (TTV) technique is an efficient and powerful tool to detect small, unseen companions of already discovered exoplanets (see Figure \ref{exoplanets}). Kepler provided the firsts breaktrough discoveries of exoplanets showing TTV, and after the mission ended some project started to complete its legacy searching for TTVs. The Transit Monitoring in the South project (TraMoS) is one of its kind in the southern hemisphere, aiming to detect multi-exoplanetary systems trough the TTV method.

\begin{figure}[t]
\centering
\includegraphics[width=1.0\columnwidth]{imagenes/exoplanets_hd.pdf}
\caption{Mass and orbital period distribution of Exoplanets, and their corresponding discovery methods in color. This plot only shows the exoplanets with known masses}
\label{exoplanets}
\end{figure}

@@

This thesis is organized as follows: Chapter \ref{chap:tramos} summarizes the TraMoS project, its goals, techniques and achievements to date. Also, this chapter includes what we can learn from transit exoplanets' light curves and Transit Timing Variations. Chapter \ref{chap:paper} shows the analysis and results of the hot Jupiters WASP-18b, WASP-19b and WASP-77Ab, including  their observations and data reduction, light curve and RV analysis, measurements of TTVs and upper mass limit for possible companions in their systems. Then, Chapter \ref{chap:tess}  recapitulates the current stage of the TraMoS project on which TESS data is being incorporated. Here, I re-analyze the TTV curves of the previous targets WASP-18b, WASP-19b, and WASP-77Ab with new light curves from TESS. Finally, Chapter \ref{chap:conclusion} contains the principal conclusions of this thesis as well as some ideas for future work. The Appendix section includes a brief description of the aperture photometry pipeline utilized, all the Python scripts used to generate each light curve, and an example of the PDF file obtained after the aperture photometry procedure.

>>>>

Having billions of billions of galaxies in the Universe each one with millions of stars, the current number of discovered exoplanets is still very low. Furthermore, the hunting zone of exoplanets is limited by our current technology and finding an exoplanet in the closer galaxy -- Andromeda -- could be considered science fiction. If the Milky Way has \textit{only} around $2\,\times10^{11}$ stars, how could be the Solar System unique? How could be the Earth the only planet with life on it? Until today, it is still difficult to answer those questions. The Solar System presents a logic structure where the small, rocky planets are closer to the Sun, and the gas giants are located beyond. But no other exoplanetary system have been found having similar features and even not exoplanetary life has been detected yet\footnote{By the date of writing this thesis.}. Our knowledge about the Solar System challenges the existence of other kinds of planetary architectures with close-in gas giants, for instance, suggesting that there is not a single, universal planetary process formation. Answering such questions makes studying exoplanets a relevant matter nowadays.

@@

Anyway, the rising research field of exoplanets does not end with their discovery. Performing follow-up and proper characterization of discovered exoplanets is essential to obtain a full list of orbital and planetary parameters. For instance,  39\%  of the total number of discovered exoplanets have a proper measurement of their masses, and only 17\% have their mass and radius computed. Moreover, through follow-up studies several phenomena, such as their atmospheres, can be unveiled.

\begin{figure}[t]
\centering
\includegraphics[width=1.0\columnwidth]{imagenes/exoplanets_hd.pdf}
\caption{Mass and orbital period distribution of Exoplanets, and their corresponding discovery methods in color. This plot only shows the exoplanets with measured masses. The Transit and Radial Velocity methods (in orange and blue, respectively) are the most successful techniques to discover extrasolar planets. However, both techniques are still more sensitive to detect Jupiter-mass planets rather than Earth-mass, but the Transit method is more sensitive to short-period planets.}
\label{exoplanets}
\end{figure}

Photometric follow-up of transiting exoplanets provides not only fundamental information about the exoplanet but their environment. Variations in the transit timing -- when the planet passes in front of its star -- could be induced by companions gravitationally bounded, such as other planets in their systems or even exomoons \citep{Kipping2009a,Kipping2009b}. As most of the space mission have strict fields of view, performing photometric follow-up with ground-based telescopes is crucial to reveal important features of their systems. The Transit Timing Variation (TTV) technique \citep{Holman2005,Agol2005} is an efficient and powerful tool to detect small, unseen companions of already discovered exoplanets (see Figure \ref{exoplanets}). Kepler provided the firsts breakthrough discoveries of exoplanets showing TTV \citep{Holman2010}, and after the mission ended some projects started aiming to complete its legacy searching for TTVs. The Transit Monitoring in the South project (TraMoS) is one of its kind in the southern hemisphere, aiming to detect multi-exoplanetary systems trough the TTV method.

@@

This thesis is organized as follows: Chapter \ref{chap:tramos} summarizes the TraMoS project, its goals, techniques and achievements to date. Also, this chapter includes what we can learn from transit exoplanets' light curves and Transit Timing Variations. Chapter \ref{chap:paper} shows the analysis and results of the hot Jupiters WASP-18b, WASP-19b and WASP-77Ab, including  their observations and data reduction, light curve and RV analysis, measurements of TTVs and upper mass limit for possible companions in their systems. Then, Chapter \ref{chap:tess}  recapitulates the current stage of the TraMoS project on which TESS data is being incorporated. Here, I re-analyze the TTV curves of the previous targets WASP-18b, WASP-19b, and WASP-77Ab with new light curves from TESS. Finally, Chapter \ref{chap:conclusion} contains a summary of this thesis as well as some ideas for future work. The Appendix section includes a brief description of the aperture photometry pipeline utilized, all the Python scripts used to generate each light curve, and an example of the PDF file obtained after the aperture photometry procedure.

### 2019-08-19 tess.tex
TraMoS chapter ready.
Final draft.

Thus, for the currently and future stage of the TraMoS project, archival data - collected since 2008 - will be combined with the upcoming TESS data and new ground-based observations using one-meter class telescopes. The main goal of the project will remain the same: refine orbital and physical parameters of exoplanetary system and search for variations in the transit time, that could suggest the presence of additional small bodies in the system.  The KOINet project proved that ground-based follow-up are, in some cases, essential to complete a possible TTV signal, and therefore, to confirm or rule out variations in the transit time. During its first year of mission, TESS will supply with consecutive transit light curves from southern targets, specially from short-period planets, such as hot Jupiters. This special feature is fundamental to construct TTV curves and then, complete them from ground-based follow-up.

In the following subsections, I present a re-analysis of the TTV curves of the hot Jupiters: WASP-18b, WASP-19b and WASP-77Ab including new transit times from TESS: 45, 28, and 15 for WASP-18b, WASP-19b and WASP-77Ab respectively. In order to compute the transit time of each light curve, I used EXOFASTv2 but only to fit the transit time and the flux baseline. The remaining planetary parameters were fixed with the values from Tables \ref{wasp18}, \ref{tab:wasp19}, \ref{tab:wasp77}.

@@

\section{WASP-36b}

The exoplanet WASP-36b was selected as the

>>>>

Thus, for the currently and future stage of the TraMoS project, archival data -- collected since 2008 -- will be combined with the upcoming TESS data and new ground-based observations using one-meter class telescopes. The main goal of the project will remain the same: refine orbital and physical parameters of exoplanetary system and search for variations in the transit time, that could suggest the presence of additional small bodies in the system.  The KOINet project proved that ground-based follow-up are, in some cases, essential to complete a possible TTV signal, and therefore, to confirm or rule out variations in the transit time. During its first year of mission, TESS will supply with consecutive transit light curves from southern targets, specially from short-period planets, such as hot Jupiters. This special feature is fundamental to construct TTV curves and then, complete them from ground-based follow-up.

In the following subsections, I present a re-analysis of the TTV curves of the hot Jupiters: WASP-18b, WASP-19b and WASP-77Ab including new transit times from TESS: 45, 28, and 15 for WASP-18b, WASP-19b and WASP-77Ab respectively. In order to compute the transit time of each light curve, I used EXOFASTv2 to only fit the transit time and the flux's baseline. The remaining planetary parameters were fixed with the values from Tables \ref{wasp18}, \ref{tab:wasp19}, \ref{tab:wasp77}.

@@

%\section{WASP-36b}

%The exoplanet WASP-36b was selected as the

### 2019-08-19 tramos.tex
Final draft.

The Transit Monitoring in the South (TraMoS) project started in 2008 aiming to perform photometric follow-up of transiting exoplanets using telescopes located in Chile.  Through photometric monitoring of transiting events, important information of the planetary system can be obtained, such as the planetary radius, inclination of the orbit, precise planetary mass, among others. Moreover, this kind of analysis allows the detection of variability in the transit's parameters (TPV), if several epochs of the transit event are included. The TraMoS project aims, on the one hand, to refine the physical and orbital parameters of selected exoplanets systems through photometric follow-up and, on the other hand, to search variations in that parameters that could suggest the presence of additional bodies in the system.

@@

\begin{figure}[ht]
\centering
\includegraphics[width=0.9\columnwidth]{imagenes/tramos.png}
\caption{Histogram of the 30 transiting exoplanets with the largest amount of observations within the TraMoS project. The exoplanets WASP-4b and WASP5-b were presented in \cite{Hoyer2012} and \cite{Hoyer2013}. The results from the analysis of WASP-18b, WASP-19b and WASP77-Ab are in Cortés-Zuleta et al. 2019 (submitted, see Chapter \ref{chap:paper} for further details). In none of those systems a significant TTV signal was detected.}
\label{tramos}
\end{figure}

@@

In the following section \ref{lightcurve}, I describe the light curve produced from a transiting exoplanet --the principal target of the TraMoS project --  what information about the planet can be obtain from them and how these light curves can give us information about the architecture of this kind of systems. Then, in section \ref{ttv}, I detail on an specific type of phenomena of transiting exoplanets: the variation of the transit time of its light curve (TTV).

@@

The light curve produced by a transit event provides important information of the planetary system. For example, it is the only way on which we can derive directly the proportion between the exoplanet and star radius. It can also gives information about the impact parameter, and therefore, the inclination of the orbit.

In Figure~\ref{transit} is described the architecture of a transiting exoplanet. Thanks to its geometry, the transit event as well as some parameters of the planetary system can be described with a few simple relations \cite{Seager2003,Winn2010}.  When the planet passes in front of its star produced a dimming in its flux called primary transit. The maximum drop during the transit event defines the transit depth $\Delta{F}$:

@@

\begin{figure}[H]
\centering
\includegraphics[width=1.0\columnwidth]{imagenes/transit_architecture.pdf}
\caption{Architecture of a transiting exoplanet and the observed light curve. Based on the light curve itself it is possible to measure the transit depth $\Delta F$, the impact parameter $b$. From them, the planet-to-star radius ratio $R_{p}/R_{*}$ and the orbital inclination $i$ can be derived. Figure from The Handbook of Exoplanets by M. Perryman.}
\label{transit}
\end{figure}

@@

Since the discovery of the first exoplanet, one of the principal questions that arose was: are they alone in their systems? The idea of multiplanetary systems is a consequence of our knowledge about the planetary system where we live in, and until today the Solar System has the larger number of orbiting planets. Besides the Solar System, Kepler-90 has the same number of confirmed exoplanets: eight. Statistically, around a 20\% of the confirmed exoplanets live in a multiplanetary system\footnote{Information obtained from the NASA Exoplanet Archive \url{exoplanetarchive.ipac.caltech.edu}}, but this number is theoretically larger. \cite{Fressin2013} computed the occurrence of exoplanets through numerical simulation in Kepler data, obtaining that around $52\%$ of the stars should have at least one planet within a 85 days-orbit.

@@

\cite{Holman2010} presented the first multiplanetary system confirmed using the TTV method: two planets orbiting Kepler-9, a Sun-like star. These two Saturn-size planets are in near 2:1 orbital resonance with periods of 38.9 and 19.2 days. In order to model the dynamical interaction of the system, only 9 transits of Kepler-9b and 6 from Kepler-9c were required. This confirmed the effectiveness of the TTV method to characterize planetary systems with a scarce number of transits.

Since the discovery of \cite{Holman2010}, around 20 exoplanets have been discovered and characterized thanks to the measurement of variations on their transit times and the following proper dynamical model ling of the system. Most of them are Kepler systems because of its high duty-cycle, on which it provided years on transits data. Kepler stopped working on 2018, but the Transit Exoplanet Satellite Survey (TESS) is currently providing its first results providing planetary systems showing TTVs.  For example, two exoplanets (possibly one Jupiter and a sub-Saturn like) were recently discovered by \cite{Dawson2019} orbiting TOI-216, in near 2:1 resonance.

@@

\cite{Steffen2012b} analyzed Kepler data in order to constraint the occurrence rate of companions in hot Jupiter system. In a sample of 63 hot Jupiter candidates, none of them show evidence of possible companions in their system. When comparing with other types of exoplanets, 30\% of hot Neptunes candidates could have companions and 10\% of warm Jupiter candidates. This could suggest that hot Jupiter may have a different formation processes on which smaller companions are not allowed in their system: probably they are ejected through planet-planet scatter during migration. TTV studies could provide important clues about formation processes and dynamical evolution, and their mass and period dependence.

>>>>

The Transit Monitoring in the South (TraMoS) project started in 2008 aiming to perform a photometric follow-up of transiting exoplanets using telescopes located in Chile.  Through photometric monitoring of transiting events, important information of the planetary system can be obtained, such as the planetary radius, inclination of the orbit, precise planetary mass, among others. Moreover, this kind of analysis allows the detection of variability in the transit's parameters (TPV), if several epochs of the transit event are included. The TraMoS project aims, on one hand, to refine the physical and orbital parameters of selected exoplanets systems through photometric follow-up and, on the other, to search variations in those parameters that could suggest the presence of additional bodies in the system.

@@

\begin{figure}[ht]
\centering
\includegraphics[width=0.9\columnwidth]{imagenes/tramos.png}
\caption{Histogram of the 30 transiting exoplanets with the largest amount of observations within the TraMoS project. The exoplanets WASP-4b and WASP5-b were presented in \cite{Hoyer2012} and \cite{Hoyer2013}, respectively. The results from the analysis of WASP-18b, WASP-19b and WASP77-Ab are in Cortés-Zuleta et al. 2019 (submitted, see Chapter \ref{chap:paper} for further details). In none of those systems a significant TTV signal was detected.}
\label{tramos}
\end{figure}

@@

In the following section \ref{lightcurve}, I describe the light curve produced from a transiting exoplanet --the principal target of the TraMoS project --  what information about the planet can be obtained from them and how these light curves can give us information about the architecture of this kind of systems. Then, in section \ref{ttv}, I detail on a specific type of phenomena of transiting exoplanets: the variation of the transit time of its light curve (TTV).

@@

The light curve produced by a transit event provides important information about the planetary system. For example, it is the only way on which we can derive directly the proportion between the exoplanet and star radius. It can also give information about the impact parameter, and therefore, the inclination of the orbit.

In Figure~\ref{transit} is described the architecture of a transiting exoplanet. Thanks to its geometry, the transit event, as well as some parameters of the planetary system, can be described with a few simple relations \cite{Seager2003,Winn2010}.  When the planet passes in front of its star produced a dimming in its flux called primary transit. The maximum drop during the transit event defines the transit depth $\Delta{F}$:

@@

\begin{figure}[H]
\centering
\includegraphics[width=1.0\columnwidth]{imagenes/transit_architecture.pdf}
\caption{Architecture of a transiting exoplanet and the observed light curve. Based on the light curve itself it is possible to measure the transit depth $\Delta F$ and the impact parameter $b$. From them, the planet-to-star radius ratio $R_{p}/R_{*}$ and the orbital inclination $i$ can be derived. Figure from The Handbook of Exoplanets by M. Perryman.}
\label{transit}
\end{figure}

@@

Since the discovery of the first exoplanet, one of the principal questions that arose was: are they alone in their systems? The idea of multiplanetary systems is a consequence of our knowledge about the planetary system where we live in, and until today the Solar System has the larger number of orbiting planets. Besides the Solar System, Kepler-90 has the same number of confirmed exoplanets: eight. Statistically, around a 20\% of the confirmed exoplanets live in a multiplanetary system\footnote{Information obtained from the NASA Exoplanet Archive \url{exoplanetarchive.ipac.caltech.edu}}, but this number is theoretically larger. \cite{Fressin2013} computed the occurrence of exoplanets through numerical simulation in Kepler data, obtaining that around $52\%$ of the stars should have at least one planet within a 85 days orbit.

@@

\cite{Holman2010} presented the first multi-planetary system confirmed using the TTV method: two planets orbiting Kepler-9, a Sun-like star. These two Saturn-size planets are in near 2:1 orbital resonance with periods of 38.9 and 19.2 days. To model the dynamical interaction of the system, only 9 transits of Kepler-9b and 6 from Kepler-9c were required. This confirmed the effectiveness of the TTV method to characterize planetary systems with a scarce number of transits.

Since the discovery of \cite{Holman2010}, around 20 exoplanets have been discovered and characterized thanks to the measurement of variations on their transit times and the following proper dynamical modeling of the system. Most of them are Kepler systems because of its high duty-cycle, on which it provided years on transits data. Kepler stopped working on 2018, but the Transit Exoplanet Satellite Survey (TESS) is currently delivering its first results with new planetary systems showing TTVs.  For example, two exoplanets -- possibly one Jupiter and a sub-Saturn like -- were recently discovered by \cite{Dawson2019} orbiting TOI-216, in near 2:1 resonance.

@@

\cite{Steffen2012b} analyzed Kepler data to constraint the occurrence rate of companions in hot Jupiter system. In a sample of 63 hot Jupiter candidates, none of them show evidence of possible companions in their system. When comparing with other types of exoplanets, 30\% of hot Neptune candidates may have companions and 10\% of warm Jupiter candidates. This could suggest that hot Jupiter may have different formation processes on which smaller companions are not allowed in their system: probably they are ejected through planet-planet scatter during migration. TTV studies could provide important clues about formation processes and dynamical evolution, and their mass and period dependence.
